Interval Probability for Fuzzy Quantum Theories

### 2016-06-30 prob.tex
prob.tex Write Radon-Nikodym theorem in quantum probability space...

Although specifying a probability for every event looks complex, a
probability measure can be simply constructed by $\pmeas(E)=\tensor*[_{\mathbb{R}}]{\sum}{_{\omega\in E}}\rho(\omega)$,
where $\rho:\Omega\rightarrow[0,1]$ and $\tensor*[_{\mathbb{R}}]{\sum}{_{\omega\in\Omega}}\rho(\omega)=1$.
Furthermore, every probability measure~$\pmeas$ can be written in
this form for an unique $\rho$~\cite{Kolmogorov1950,Swart2013}
which is called the probability distribution~\cite{GrahamKnuthPatashnik1994}.

@@

\begin{example}[Two-coin experiment with interval probability]
  \label{ex3} We split the unit interval $[0,1]$ in the following four
  closed sub-intervals: $[0,0]$ which we call \imposs,
  $[0,\frac{1}{2}]$ which we call \unlikely, $[\frac{1}{2},1]$ which
  we call \likely, and $[1,1]$ which we call \necess. Using these new
  values, we can modify the probability measure of Ex.~\ref{ex1} by
  mapping each numeric value to the smallest sub-interval containing
  it to get the following:
\[
\begin{array}{c@{\qquad\qquad}c}
\begin{array}{rcl}
\bar{\mu}(\emptyset) & = & \imposs\\
\bar{\mu}(\{HH\}) & = & \unlikely\\
\bar{\mu}(\{HT\}) & = & \imposs\\
\bar{\mu}(\{TH\}) & = & \likely\\
\bar{\mu}(\{TT\}) & = & \imposs\\
\bar{\mu}(\{HH,HT\}) & = & \unlikely\\
\bar{\mu}(\{HH,TH\}) & = & \necess\\
\bar{\mu}(\{HH,TT\}) & = & \unlikely
\end{array} & \begin{array}{rcl}
\bar{\mu}(\{HT,TH\}) & = & \likely\\
\bar{\mu}(\{HT,TT\}) & = & \imposs\\
\bar{\mu}(\{TH,TT\}) & = & \likely\\
\bar{\mu}(\{HH,HT,TH\}) & = & \necess\\
\bar{\mu}(\{HH,HT,TT\}) & = & \unlikely\\
\bar{\mu}(\{HH,TH,TT\}) & = & \necess\\
\bar{\mu}(\{HT,TH,TT\}) & = & \likely\\
\bar{\mu}(\{HH,HT,TH,TT\}) & = & \necess
\end{array}\end{array}
\]
Despite the absence of any numeric information, the probability
measure is quite informative: it reveals that the second coin is
double-headed and that the first coin is biased. To understand the
$\subseteq$-condition, consider the following calculation:
\begin{eqnarray*}
 &  & \bar{\mu}(\{HH\})+\bar{\mu}(\{HT\})+\bar{\mu}(\{TH\})+\bar{\mu}(\{TT\})\\
 & = & \imposs+\unlikely+\imposs+\likely\\
 & = & \left[0,0\right]+\left[0,\frac{1}{2}\right]+\left[0,0\right]+\left[\frac{1}{2},1\right]=\left[\frac{1}{2},\frac{3}{2}\right]
\end{eqnarray*}
If we were to equate $\bar{\mu}(\Omega)$ with the sum of the individual
probabilities we would get that
$\bar{\mu}(\Omega) = \left[\frac{1}{2},\frac{3}{2}\right]$. However,
using the fact that $\bar{\mu}(\emptyset)=\imposs$, we have
$\bar{\mu}\left(\Omega\right)=1-\bar{\mu}\left(\emptyset\right)=\necess=[1,1]$. This
interval is tighter and a better estimate for the probability of the
event $\Omega$ and of course it is contained in
$[\frac{1}{2},\frac{3}{2}]$. However it is only possible to exploit
the information about the complement when all four events are
combined. Thus the $\subseteq$-condition allows us to get an estimate
for the combined event from each of its constituents and then gather
more evidence knowing the aggregate event.\qed\end{example}

@@

Another difference between classical and quantum probability space. For a
given set of events $\events$, there are many possible probability measures $\mu:\events\rightarrow[0,1]$. The Born rule~\cite{Born1983,Mermin2007,RiederSvozil2007},
a postulate of quantum mechanics, states that each pure normalized
($\ip{\phi}{\phi}=1$) quantum state $\ket{\phi}$ induces a probability measure
$\mu^B_{\phi}$ as follows: 
\[
\mu^B_{\phi}(P)=\ip{\phi}{P\phi}
\]

Moreover, the Born rule can be extended to a mixed state. Suppose
we want to prepare a quantum system. As we discussed in classical
probability, our ability of preparing a state might not be perfect.
If we want to prepare $\ket{\phi}$, we may turn out preparing a set
of state~$\ket{\phi_{j}}$ each with probability~$q_{j}$. Then,
the state of the system can be expressed as a density matrix~$\rho=\sum_{j}q_{j}\proj{\phi_{j}}$,
where $\sum_{j}q_{j}=1$. It is natural that the quantum probability
measure introduced by $\rho$ is the combination of $\mu^B_{\phi_{j}}$
with respect to probability $q_{j}$~\cite{peres1995quantum,544199,RiederSvozil2007}:
\begin{eqnarray}
\mu^B_{\rho}\left(P\right) & = & \Tr\left(\rho P\right)=\sum_{j=1}^{N}q_{j}\mu^B_{\phi_{j}}\left(P\right)\textrm{ .}\label{BornRule.mixed}
\end{eqnarray}

Conversely, in Hilbert spaces of dimension $d\geq3$, Gleason's theorem
states that given a probability measure $\mu$, there exist a mixed
state~$\rho$ that induces such a measure using the Born rule~\cite{gleason1957,Redhead1987-REDINA,peres1995quantum}.
By applying Gleason's theorem, we can extend lemma~\ref{lem:quantumProbabilityMeasure}
to the following corollary.

>>>>

Although specifying a probability for every event looks complex, a
probability measure can be simply constructed by $\pmeas(E)=\tensor*[_{\mathbb{R}}]{\sum}{_{\omega\in E}}\rho(\omega)$,
where $\rho:\Omega\rightarrow[0,1]$ and $\tensor*[_{\mathbb{R}}]{\sum}{_{\omega\in\Omega}}\rho(\omega)=1$.
Furthermore, every probability measure~$\pmeas$ can be written in
this form for an unique $\rho$~\cite{Kolmogorov1950,Swart2013}.
This fact is a special case of the the Radon-Nikodym theorem, and
$\rho$ is called the Radon-Nikodym derivative of $\pmeas$ with respect
to the counting measure~\cite{Nikodym1930,Folland2013}. For example,
the Radon-Nikodym derivative of $\pmeas$ in example~\ref{ex1} is:
\[
\begin{array}{rcl}
\rho\left(HH\right)=1/3,\quad\rho\left(HT\right)=0,\quad\rho\left(TH\right)=2/3,\quad\rho\left(TT\right)=0\textrm{ .}\end{array}
\]

@@

\begin{example}[Two-coin experiment with interval probability]
  \label{ex3} We split the unit interval $[0,1]$ in the following four
  closed sub-intervals: $[0,0]$ which we call \imposs,
  $[0,\frac{1}{2}]$ which we call \unlikely, $[\frac{1}{2},1]$ which
  we call \likely, and $[1,1]$ which we call \necess. Using these new
  values, we can modify the probability measure of Ex.~\ref{ex1} by
  mapping each numeric value to the smallest sub-interval containing
  it to get the following:
\[
\begin{array}{c@{\qquad\qquad}c}
\begin{array}{rcl}
\bar{\mu}(\emptyset) & = & \imposs\\
\bar{\mu}(\{HH\}) & = & \unlikely\\
\bar{\mu}(\{HT\}) & = & \imposs\\
\bar{\mu}(\{TH\}) & = & \likely\\
\bar{\mu}(\{TT\}) & = & \imposs\\
\bar{\mu}(\{HH,HT\}) & = & \unlikely\\
\bar{\mu}(\{HH,TH\}) & = & \necess\\
\bar{\mu}(\{HH,TT\}) & = & \unlikely
\end{array} & \begin{array}{rcl}
\bar{\mu}(\{HT,TH\}) & = & \likely\\
\bar{\mu}(\{HT,TT\}) & = & \imposs\\
\bar{\mu}(\{TH,TT\}) & = & \likely\\
\bar{\mu}(\{HH,HT,TH\}) & = & \necess\\
\bar{\mu}(\{HH,HT,TT\}) & = & \unlikely\\
\bar{\mu}(\{HH,TH,TT\}) & = & \necess\\
\bar{\mu}(\{HT,TH,TT\}) & = & \likely\\
\bar{\mu}(\{HH,HT,TH,TT\}) & = & \necess
\end{array}\end{array}
\]
Despite the absence of any numeric information, the probability
measure is quite informative: it reveals that the second coin is
double-headed and that the first coin is biased. To understand the
$\subseteq$-condition, consider the following calculation:
\begin{eqnarray*}
 &  & \bar{\mu}(\{HH\})+\bar{\mu}(\{HT\})+\bar{\mu}(\{TH\})+\bar{\mu}(\{TT\})\\
 & = & \imposs+\unlikely+\imposs+\likely\\
 & = & \left[0,0\right]+\left[0,\frac{1}{2}\right]+\left[0,0\right]+\left[\frac{1}{2},1\right]=\left[\frac{1}{2},\frac{3}{2}\right]
\end{eqnarray*}
If we were to equate $\bar{\mu}(\Omega)$ with the sum of the individual
probabilities we would get that
$\bar{\mu}(\Omega) = \left[\frac{1}{2},\frac{3}{2}\right]$. However,
using the fact that $\bar{\mu}(\emptyset)=\imposs$, we have
$\bar{\mu}\left(\Omega\right)=1-\bar{\mu}\left(\emptyset\right)=\necess=[1,1]$. This
interval is tighter and a better estimate for the probability of the
event $\Omega$ and of course it is contained in
$[\frac{1}{2},\frac{3}{2}]$. However it is only possible to exploit
the information about the complement when all four events are
combined. Thus the $\subseteq$-condition allows us to get an estimate
for the combined event from each of its constituents and then gather
more evidence knowing the aggregate event.\qed\end{example}
\footnote{\yutsung{The Radon-Nikodym Theorem for interval-valued
probability?\cite{GilboaSchmeidler1994}?}}

@@

Like a classical probability measure can be easily defined by its
Radon-Nikodym derivative, a quantum probability measure can be easily
constructed by states according to the Born rule~\cite{Born1983,Mermin2007,RiederSvozil2007}.
For each pure normalized ($\ip{\phi}{\phi}=1$) quantum state $\ket{\phi}$,
the Born rule induces a probability measure $\mu_{\phi}^{B}$ as follows:
\[
\mu_{\phi}^{B}(P)=\ip{\phi}{P\phi}\textrm{ .}
\]

Moreover, the Born rule can be extended to a mixed state. Suppose
we prepares a set of state~$\ket{\phi_{j}}$ each with probability~$q_{j}$,
the state of the system can be expressed as a density matrix~$\rho=\sum_{j}q_{j}\proj{\phi_{j}}$,
where $\sum_{j}q_{j}=1$. It is natural that the quantum probability
measure introduced by $\rho$ is the combination of $\mu_{\phi_{j}}^{B}$
with respect to probability $q_{j}$~\cite{peres1995quantum,544199,RiederSvozil2007}:
\begin{eqnarray}
\mu_{\rho}^{B}\left(P\right) & = & \Tr\left(\rho P\right)=\sum_{j=1}^{N}q_{j}\mu_{\phi_{j}}^{B}\left(P\right)\textrm{ .}\label{BornRule.mixed}
\end{eqnarray}

The generalized Radon-Nikodym theorem is called Gleason's theorem
which states that, in Hilbert spaces of dimension $d\geq3$, given
a probability measure $\mu$, there exist a mixed state~$\rho$ that
induces such a measure using the Born rule~\cite{gleason1957,Redhead1987-REDINA,peres1995quantum}.
By applying Gleason's theorem, we can extend lemma~\ref{lem:quantumProbabilityMeasure}
to the following corollary.
